---
title: "fars package details"
author: "Jacob Greif"
date: "2021-10-27"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{fars_details}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---
```{r  global_options}
knitr::opts_chunk$set(echo = FALSE, message = FALSE,
  warning = FALSE)

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


```{r setup, echo=FALSE, message=FALSE}
#library(devtools)
#library(dplyr)

#install_github("jakegreif/FARS", auth_token = "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")
#library(FARS)
```

# Introduction

The FARS package contains functions that read, create, and summarize files 
containing data from the US National Highway Traffic Safety Administration's 
Fatality Analysis Reporting System (FARS). FARS is a nationwide census providing
the American public yearly data regarding fatal injuries suffered in motor 
vehicle traffic crashes.

This vignette uses example data to demonstrate the package's capabilities and 
how to use these functions within a workflow.

# Example Data
The FARS data used in this vignette is from the years 2013, 2014, and 2015. You
can learn more about the FARS data set [here](https://www.nhtsa.gov/research-data/fatality-analysis-reporting-system-fars).

# Reading in FARS Data

The first step to begin using your FARS data is to coerce the data into a usable 
format in R. The "fars_read" function does this for you by converting your FARS
csv data into a `tbl_df` (tibble data frame).

```{r fars_read example}
#fars <- fars_read("accident_2014.csv.bz2")
#class(fars)
```

## Creating a File name

Additionally, if you want to generate a new file with FARS data you can use the 
"make_filename" function to generate a file name that's consistent with other
FARS data files.

```{r make_filename example}
#make_filename(2019)
```

# Examinining FARS Data

## From a Single Year

A user can examine FARS data from a single year using the "fars_read_years"
function. The function sorts the data by month and year, and it returns a list 
with a length of one. The first element element in the list is the `tbl_df` (the
`tidyverse` data frame) listing the month and year for each fatal accident.

```{r fars_read_years single year}
#fars2014 <- fars_read_years(2014)
#fars2014.df <- fars2014[[1]]
```

## From Multiple Years

Alternatively, a user can provide a list of years as the argument for 
"fars_read_years", which returns a list of `tbl_df`s where ach element contains 
the month and year for each fatality.

```{r fars_read_years mulitple years}
#fars2013_2015 <- fars_read_years(c(2013, 2014, 2015))
#fars2013_2015
```

# Summarizing FARS Data

If a user wants to learn more about their FARS data, they can use the 
"fars_summarize_years" function to see summary statistics for each month in the
selected year(s). Just like the "fars_read_years" function, the user can provide
a single year as an argument, or a list of multiple years.

```{r fars_summarize_years examples}
# Single year example
#fars_summarize_years(years = 2015)

# Multi-year example
#fars_summarize_years(years = c(2013, 2014, 2015))
```

# Mapping FARS data

One can use the "fars_map_state" function to visualize FARS data spatially for
a given state and year. The function generates a simple map (using the 'mapdata'
package, which you may need to install and/or add to your library) of the 
selected state with points representing the location of each fatality in the 
selected year.

```{r fars_map_state example}
#fars_map_state(state.num = 28, year = 2015)
```
